---
title: "GOM-series NOAA Interannual Climate Indices"
output: github_document
---

[NOAA Interannual Climate Indices](https://psl.noaa.gov/data/climateindices/list/)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
source("setup.R")
```



## Atlantic Multidecadal Oscillation (AMO) Index - long version

Enfield, D.B., A. M. Mestas-Nunez and P.J. Trimble, 2001: The Atlantic multidecadal oscillation and it's relation to rainfall and river flows in the continental U.S.. Geophysical Research Letters, Vol. 28, 2077-2080.

```{r}
amo_wide <- fetch_amo(form="wide") |>
  glimpse()
```

```{r}
amo_long <- fetch_amo(form = "long") |>
  glimpse()
```

```{r}
ggplot(amo_long, aes(x=date, y=amo)) +
  geom_line()
```


## North Atlantic Oscillation (NAO) Index

Hurrell, J.W., 1995: Decadal trends in the North Atlantic Oscillation and relationships to regional temperature and precipitation. Science 269, 676-679.

Jones, P.D., Jónsson, T. and Wheeler, D., 1997: Extension to the North Atlantic Oscillation using early instrumental pressure observations from Gibraltar and South-West Iceland. Int. J. Climatol. 17, 1433-1450.


```{r}
fetch_nao(form="wide") |>
  glimpse()
```

```{r}
nao_long <- fetch_nao(form = "long") |>
  glimpse()
```


```{r}
ggplot(nao_long, aes(x=date, y=nao)) +
  geom_line()
```


