---
title: "GOM-Series Buoys"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
suppressPackageStartupMessages({
  library(cofbb)
  library(ggOceanMaps)
})
source("setup.R")
```

## Buoy data aggregated to monthly means

5 buoys from across the Gulf of Maine were selected for inclusion.  A listing of these is saved in `data/buoy-listings.csv` and can be read as simple look-up table (lut).

```{r}
buoys = buoy_lut()
buoys
```

```{r, warning = FALSE}
bb <- cofbb::get_bb("gom")
basemap(limits = bb, bathymetry = TRUE) + 
  geom_point(data = buoys, aes(x = lon, y = lat), 
             color = "orange") +
  geom_text(data = buoys, aes(x = lon, y = lat, 
                              label = id,
                              hjust = 0.5, 
                              vjust = 1.2))
```


## MET

### Fetch meteorological (met) data
Meteorological data for these buoys can be fetched using `fetch_buoy_met()`. This function downloads high temporal resolution data per buoy, and aggregates into monthly means, and saved to disk a simple table. Run this as needed to update data.

```
met <- lapply(buoys$id, fetch_buoy_met)
```

### Read and display met data

Read one or more buoy met data files using `read_buoy_met()`.  By default all buoys are read and boundinto one table.

```{r}
x <- read_buoy_met() |>
  dplyr::mutate(month = format(date, "%b"), .after = date) |>
  dplyr::group_by(buoy)

ggplot(data = x, aes(x = date, y = wind_speed, color = buoy, shape = buoy)) +
  geom_line()
```

```{r}
ggplot(data = filter(x, month == 'Aug'), 
       aes(x = date, y = air_temperature, color = buoy)) +
  geom_point() + 
  geom_smooth(method = "lm", se=FALSE) +
  labs(x = "year", y = 'Air Temp (C)', title = "August - mean monthly air temperature")
```


## CTD at depth (temp, salinity, sigma_t)

### Fetch CTD  data
CTD data for these buoys can be fetched using `fetch_buoy_ctd()`. This function downloads high temporal resolution data per buoy at various depths, and aggregates into monthly means, and saved to disk a simple table. Run this as needed to update data.

```
ctd <- lapply(buoys$id, fetch_buoy_ctd)
```

### Read and display CTD data

Read one or more buoy CTD data files using `read_buoy_ctd()`.  By default all buoys are read and bound into one table.

```{r}
x <- read_buoy_ctd() |>
  dplyr::mutate(depth = factor(as.integer(depth))) |>
  dplyr::mutate(month = format(date, "%b"), .after = date) |>
  dplyr::group_by(buoy)

ggplot(data = x, aes(x = date, y = temperature, color = depth)) +
  scale_y_reverse()  + 
  geom_line() + 
  facet_wrap(~buoy)
```

