---
title: "GOM-series stream guage data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
source("setup.R")
```

# How to access [USGS Stream Gauge data for Maine](https://waterdata.usgs.gov/me/nwis/rt) 

## List stations of interest

```{r}
stations <- usgs_lut()
stations
```


## Use the `readNWISdv` function from the [dataRetrieval]() package to get daily discharge values from the stations of interest

The parameter code "00060" is for daily discharge values in ft^3/s

```{r}
x <- fetch_usgs(stations, parameter = "00060")
x
```


## Take a look at how much data is available per station

```{r}
ggplot(data=x, aes(x=date, y=discharge)) +
  geom_line() +
  facet_grid(rows = "site_no")
```

### Aggregate into monthly intervals


Here we select one station, "Androscoggin River near Auburn, Maine", and display the min and max monthly discharge.  Note that we have a selection of 6 metrics summarizing each month.

```{r}
andro_month = usgs_aggregate(x, by = "month") |>
  dplyr::mutate(month = format(date, "%b")) |>
  dplyr::filter(site_no == "01059000") |>
  dplyr::glimpse()
```

For the display we logscale to help reveal the "texture" of the minimum flow rate.

```{r}
ggplot(data = andro_month, aes(x = date, y = min)) +
  labs(y = "min and max log10(discharge)", title = "Androscoggin River near Auburn, Maine") + 
  geom_line(color = "blue", linewidth = 0.5) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_line(aes(x = date, y = max), color = "orange", linewidth = 0.5) + 
  geom_smooth(aes(x = date, y = max), method = "lm", se = FALSE, color = "orange") +
  scale_y_log10() + 
  scale_x_date(
    NULL,
    breaks = scales::breaks_width("25 years"),
    labels = scales::label_date("'%y")) + 
  facet_wrap(~ month)
```

We can compute the same for annual summaries.

```{r}
andro_year = usgs_aggregate(x, by = "year") |>
  dplyr::filter(site_no == "01059000") |>
  dplyr::glimpse()
```

```{r}
ggplot(data = andro_year, aes(x = date, y = q25)) + 
  geom_line(color = "blue", linewidth = 0.5) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_line(aes(y = q75), color = "orange") +
  geom_smooth(aes(x = date, y = q75), method = "lm", se = FALSE, color = "orange") +
  scale_x_date(
    NULL,
    breaks = scales::breaks_width("5 years"),
    labels = scales::label_date("'%y")) + 
  labs(y = "First and Third quartiles log10(discharge)",
       title = "Androscoggin River near Auburn, Maine") + 
  facet_wrap(~site_no)
```
