---
title: "GOM-series GHCN daily"
output: github_document
---

CANADA # https://climatedata.ca/download/#station-download
https://api.weather.gc.ca/collections/climate-daily/items?datetime=1840-03-01%2000:00:00/2023-05-02%2000:00:00&STN_ID=6244&sortby=PROVINCE_CODE,STN_ID,LOCAL_DATE&f=csv&limit=150000&startindex=0

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
source("setup.R")
```

### GHCNd

We use the [ghcnd](https://github.com/BigelowLab/ghcnd) R package to downlaod and wrangle daily data from NOAA's [Global Historical Climatology Network daily (GHCNd)](https://www.ncei.noaa.gov/products/land-based-station/global-historical-climatology-network-daily).  We aggregate these daily values into monthly means.

We need to narrow the selection of stations into those along that lie around the perimeter of the Gulf of Maine.  We also hope to find stations that provide long time series covering the intervals 1950-2000 and 2000-present.

```{r}
station_ids = c(PWM = "USW00014764",
                BHC = "USC00190736",
                #LIA = "USW00014739",
                DNH = "USC00272174",
                ANP = "USC00170100",
                COR = "USC00171628")
stations = fetch_station_list() |>
  dplyr::filter(id %in% station_ids)

inv = fetch_inventory(stations = stations) |>
  dplyr::mutate(duration = lastyear - firstyear + 1)

bb = sf::st_bbox(c(xmin = -72, ymin = 39, xmax = -63,  ymax = 46), crs = 4326) |>
  sf::st_as_sfc()
coast = rnaturalearth::ne_coastline(scale= 'medium', returnclass = 'sf') |>
  sf::st_crop(bb)

```

```{r, warning = FALSE}
tmax = filter(inv, element == "TMAX")
ggplot(data = tmax) +
  geom_sf(size = 1.5) +
  geom_sf(data = st_geometry(coast)) +
  geom_sf_text(data = tmax, aes(label = name), 
               color = "blue", hjust = -0.1)
```
We can now read in the per station per element data.  We group the data into two epochs, 1950-1999 and 2000-present.
```{r}
DATES = c(as.Date(c("1950-01-01", "2000-01-01")), Sys.Date())
x = fetch_station(inv) |>
  aggregate_monthly() |>
  dplyr::filter(DATE >= DATES[1]) |>
  dplyr::mutate(TDIFF = TMAX - TMIN, .before = geometry)
ix <- findInterval(x$DATE, DATES)
x = dplyr::ungroup(x) |>
  dplyr::mutate(EPOCH = factor(as.character(sort(unique(ix)))[ix]), .after = DATE) |>
  dplyr::group_by(STATION, EPOCH) 
```


Let's look at the station data - plotting the TMIN-to-TMAX range per month range as a gray ribbon.  Superimposed are the loess smoothing lines for each epoch.
```{r, warning=FALSE}
stations = unique(x$NAME)

plot_envelope = function(x){
  ggplot(data = x,  aes(x = DATE, y = TMIN)) +
  geom_ribbon(aes(ymin = TMIN, ymax = TMAX), fill = 'gray50') +
   geom_smooth(data = dplyr::filter(x, EPOCH == 1), 
    aes(x = DATE, y = TMIN), method = loess, formula = y ~ x, se = FALSE, color = "blue") +
  geom_smooth(data = dplyr::filter(x, EPOCH == 2), 
    aes(x = DATE, y = TMIN), method = loess, formula = y ~ x, se = FALSE, color = "orange") +
  geom_smooth(data = dplyr::filter(x, EPOCH == 1), 
    aes(x = DATE, y = TMAX), method = loess, formula = y ~ x, se = FALSE, color = "blue") +
  geom_smooth(data = dplyr::filter(x, EPOCH == 2), 
    aes(x = DATE, y = TMAX), method = loess, formula = y ~ x, se = FALSE, color = "orange") +
  labs(x = "Date", y = "Minimum Monthly Temperature (C)", title = x$NAME[1]) + 
  facet_wrap(~ MONTH, scales = "free_y") 
}

gg = lapply(stations, function(name) plot_envelope(dplyr::filter(x, NAME == name)))

for (g in gg) print(g)
```

Next we can view anomlay heat maps.

```{r, fig.asp = 0.9}
for (station in stations) {
  plot( stsaav::stsaav(dplyr::filter(x, NAME == station),
                                     t_step = "Month",
                                     tcol = "DATE",
                                     vcol = "TMIN"),
        main = paste0(station, ", Monthly Mean Temp (C)"))
}
```