---
title: "GOM-series NOAA Interannual Climate Indices"
output: github_document
---

[NOAA Interannual Climate Indices](https://psl.noaa.gov/data/climateindices/list/)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
source("../setup.R")
```



## Atlantic Multidecadal Oscillation (AMO) Index - long version

Enfield, D.B., A. M. Mestas-Nunez and P.J. Trimble, 2001: The Atlantic multidecadal oscillation and it's relation to rainfall and river flows in the continental U.S.. Geophysical Research Letters, Vol. 28, 2077-2080.

```{r}
read_amo(form="wide") |>
  glimpse()
```

```{r}
amo_long <- read_amo(form = "long") |>
  glimpse()
```

```{r}
ggplot(amo_long, aes(x=date, y=amo)) +
  geom_line()
```


## North Atlantic Oscillation (NAO) Index

Hurrell, J.W., 1995: Decadal trends in the North Atlantic Oscillation and relationships to regional temperature and precipitation. Science 269, 676-679.

Jones, P.D., Jónsson, T. and Wheeler, D., 1997: Extension to the North Atlantic Oscillation using early instrumental pressure observations from Gibraltar and South-West Iceland. Int. J. Climatol. 17, 1433-1450.


```{r}
read_nao(form="wide") |>
  glimpse()
```

```{r}
nao_long <- read_nao(form = "long") |>
  glimpse()
```


```{r}
ggplot(nao_long, aes(x=date, y=nao)) +
  geom_line()
```


## Gulf Stream Index (GSI)


```{r}
read_gsi() |>
  glimpse()
```


Alternatively, one can read all three at once in either wide or long format.

```{r}
read_climate_indices() |>
  glimpse()

read_climate_indices(form = "long") |>
  glimpse()
```


# Aggregating and exporting

## Aggregate a climate index by year

```{r}
year = aggregate_climate_indices(by = 'year') |>
  glimpse()
```

We can retrieve the same but in long format, which is handy to ggplot style plotting.

```{r}
year = aggregate_climate_indices(by = 'year', form = 'long') |>
  glimpse()
```
```{r}
ggplot(data = year |> filter(measure == 'median'),
       aes(x = date, y = value, group = index)) +
  geom_line() +
  geom_line(data = year |> filter(measure == 'min'), col = 'blue') + 
  geom_line(data = year |> filter(measure == 'max'), col = 'orange') +
  facet_wrap( ~ index)
```