---
title: "GOM-series sea surface temperature"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
source("../setup.R")
```

### OISST monthly means

We downloaded monthly mean SST data from [OISST](https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.highres.html).  For each region we computed the pixel statistics values at each month.

> Note that we store the OISST data for by date for each subregion.  This file can be updated as needed using `fetch_oisst()` - see below.

```{r, message = FALSE}
x <- read_oisst() |> 
  dplyr::mutate(month = factor(format(date, "%b"), levels = month.abb), .after = 1) |>
  dplyr::group_by(region) |>
  glimpse()
```

```{r, message = FALSE}
ggplot(data = x, aes(x = date, y = mean)) +
  geom_line() + 
  labs(y = "Mean SST") + 
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ region)
```

```{r, message = FALSE}
georges_basin = dplyr::filter(x, region == 'Georges Basin')
ggplot(data = georges_basin, aes(x = date, y = min)) +
  labs(y = "Min and Max SST") + 
  geom_line() + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_line(aes(x = date, y = max)) + 
  geom_smooth(aes(x = date, y = max), method = "lm", se = FALSE, color = "orange") +
  facet_wrap(~ month)
```

Another monthly analysis plot.

```{r}
name = "Georges Basin"
plot(stsaav::stsaav(georges_basin,
                     t_step = "Month",
                     tcol = "date",
                     vcol = "mean"),
    main = paste0(name, ", Monthly Mean Temp (C)"))
```


### Annual Metrics

Annual means can be computed from the various sst metrics since each month has the same number of pixels are sampled for each region. Partial years are not computed.

Note that we compute only the following... the min of the min, the median of the median, the mean of the mean and the max of the max.  Then we plot the max and min by region.

```{r, message = FALSE}
a <- aggregate_oisst(x) |>
  dplyr::group_by(region)

ggplot(data = a, aes(x = date, y = median)) + 
  geom_line(color = "black", alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  geom_line(aes(x = date, y = min), color = "blue", alpha = 0.7) +
  geom_smooth(aes(x = date, y = min), method = "lm", se = FALSE, color = "blue") +
  geom_line(aes(x = date, y = max), color = "orange", alpha = 0.7) + 
  geom_smooth(aes(x = date, y = max), method = "lm", se = FALSE, color = "orange") + 
  labs(y = "Median, Min and Max SST") + 
  facet_wrap(~region)
```


### Updating the local dataset

It's easy to build or renew the local dataset.

```
r <- fetch_oisst()
```