---
title: "Correlation Plot Comparison"
output: github_document
---


```{r message=FALSE}
source("../setup.R")
```


## Read analysis data from the export file

```{r}
x = read_export(by = 'year', 
                selection = read_target_vars(treatment = c("median")),
                replace_names = TRUE, 
                standardize = FALSE) |>
  dplyr::filter(date >= as.Date("1900-01-01"))

glimpse(x)
```

Let's also read the table with standardized values

```{r}
x_standard = read_export(by = 'year', 
                         selection = read_target_vars(treatment = c("median")),
                         replace_names = TRUE, 
                         standardize = TRUE) |>
  dplyr::filter(date >= as.Date("1900-01-01"))

glimpse(x_standard)
```

## Correlation test using 1900-2022

Here's the correlation plot for the non-standardized (median) export table using 1900 to present and 'pairwise.complete.obs'

Note that we are using `corrr::rearrange()` which arranges the matrix based on correlations, so if the order changes, the test is yielding a different result. 

```{r}
c <- x |>
  correlate() |>
  rearrange() |>
  shave()

rplot(c) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


Correlation plot for the standardized version yields an almost identical result

```{r message=FALSE, warning=FALSE}
c <- x_standard |>
  correlate() |>
  rearrange() |>
  shave()

rplot(c) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


## What if we use 'complete.obs'?

The plot changes. We see different relationships come out in the plot and the variables are ordered differently on each axis.

```{r message=FALSE}
c <- x_standard |>
  correlate(use="complete.obs") |>
  rearrange() |>
  shave()

rplot(c) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


## Correlation test using 1970-2022

```{r}
x_standard <- x_standard |>
  filter(date >= as.Date("1970-01-01"))
```

Standardized version starting in 1970 still using 'pairwise.complete.obs'

```{r message=FALSE, warning=FALSE}
c <- x_standard |>
  correlate() |>
  rearrange() |>
  shave()

rplot(c) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Standardized version staring in 1970 using 'complete.obs'

```{r message=FALSE, warning=FALSE}
c <- x_standard |>
  correlate(use="complete.obs") |>
  rearrange() |>
  shave()

rplot(c) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

So, the question is: which do we choose? Or maybe: what information do we want the reader to gain from this figure?