---
title: "GOM-series GHCN daily"
output: github_document
---

> For Canadian data see [ClimateData](https://climatedata.ca) and the [inventory](https://climatedata.ca/download/#station-download). 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
source("../setup.R")
```

### GHCNd

We use the [ghcnd](https://github.com/BigelowLab/ghcnd) R package to downlaod and wrangle daily data from NOAA's [Global Historical Climatology Network daily (GHCNd)](https://www.ncei.noaa.gov/products/land-based-station/global-historical-climatology-network-daily).  We aggregate these daily values into monthly means.

We need to narrow the selection of stations into those along that lie around the perimeter of the Gulf of Maine.  We also hope to find stations that provide long time series covering the intervals 1950-2000 and 2000-present.  We retrieved the sation listing and saved it for subsequent use.  This code block is disabled so the file is not rewritten needlessly.

```
station_ids = c(PWM = "USW00014764",
                BHC = "USC00190736",
                #LIA = "USW00014739",
                DNH = "USC00272174",
                ANP = "USC00170100",
                COR = "USC00171628")
stations = fetch_station_list() |>
  dplyr::filter(id %in% station_ids) |>
  sf::write_sf(here::here("data","ghcn", "station_list.gpkg"))
```

For subsequent use you can use `ghcn_lut()` to read it form the local file.

```{r}
stations = ghcn_lut()
stations
```

Now we can access a more detailed inventory listing for each station.

```{r}
inv = fetch_inventory(stations = stations) |>
  dplyr::mutate(duration = lastyear - firstyear + 1)
inv
```

And show the locations of the stations.

```{r, warning = FALSE}
bb = sf::st_bbox(c(xmin = -72, ymin = 39, xmax = -63,  ymax = 46), crs = 4326) |>
  sf::st_as_sfc()
coast = rnaturalearth::ne_coastline(scale= 'medium', returnclass = 'sf') |>
  sf::st_crop(bb)

tmax = filter(inv, element == "TMAX")
ggplot(data = tmax) +
  geom_sf(size = 1.5) +
  geom_sf(data = st_geometry(coast)) +
  geom_sf_text(data = tmax, aes(label = name), 
               color = "blue", hjust = -0.1)
```
We can now read in the per station per element data.  We group the data into two epochs, 1950-1999 and 2000-present.
```{r}
DATES = c(as.Date(c("1950-01-01", "2000-01-01")), Sys.Date())
x = fetch_station(inv) |>
  dplyr::filter(DATE >= DATES[1]) |>
  dplyr::mutate(TDIFF = TMAX - TMIN, 
                MONTH = format(DATE, "%b"), 
                .before = geometry)
ix <- findInterval(x$DATE, DATES)
x = dplyr::ungroup(x) |>
  dplyr::mutate(EPOCH = factor(as.character(sort(unique(ix)))[ix]), .after = DATE) |>
  dplyr::group_by(STATION, EPOCH) 
```


Let's look at the station data - plotting the TMIN-to-TMAX range per month range as a gray ribbon.  Superimposed are the loess smoothing lines for each epoch.
```{r, warning=FALSE}
plot_envelope = function(x){
  ggplot(data = x,  aes(x = DATE, y = TMIN)) +
  geom_ribbon(aes(ymin = TMIN, ymax = TMAX), fill = 'gray50') +
   geom_smooth(data = dplyr::filter(x, EPOCH == 1), 
    aes(x = DATE, y = TMIN), method = loess, formula = y ~ x, se = FALSE, color = "blue") +
  geom_smooth(data = dplyr::filter(x, EPOCH == 2), 
    aes(x = DATE, y = TMIN), method = loess, formula = y ~ x, se = FALSE, color = "orange") +
  geom_smooth(data = dplyr::filter(x, EPOCH == 1), 
    aes(x = DATE, y = TMAX), method = loess, formula = y ~ x, se = FALSE, color = "blue") +
  geom_smooth(data = dplyr::filter(x, EPOCH == 2), 
    aes(x = DATE, y = TMAX), method = loess, formula = y ~ x, se = FALSE, color = "orange") +
  labs(x = "Date", y = "Minimum Monthly Temperature (C)", title = x$NAME[1]) + 
  facet_wrap(~ MONTH, scales = "free_y") 
}

gg = lapply(unique(x$STATION), function(id) plot_envelope(dplyr::filter(x, STATION == id)))

for (g in gg) print(g)
```

Next we can view anomaly heat maps, but we start with the daily data.

```{r}
x = fetch_station(inv) |>
  dplyr::filter(DATE >= DATES[1]) |>
  dplyr::mutate(TDIFF = TMAX - TMIN, .before = geometry)
```

```{r, fig.asp = 0.9}
for (id in unique(x$STATION)) {
  name = stations$name[stations$id == id]
  plot( stsaav::stsaav(dplyr::filter(x, STATION == id),
                                     t_step = "Month",
                                     tcol = "DATE",
                                     vcol = "TMIN"),
        main = paste0(name, ", Monthly Mean Temp (C)"))
}
```


An alternative is to compute intervals means for months or years.  

```{r}
a = fetch_station(inv) |>
  aggregate_ghcn(by = 'year') |>
  dplyr::mutate(YEAR = as.numeric(format(date, "%Y")))

ggplot(data = a, aes(x = YEAR, y= PRCP.mean)) + 
  geom_line() + 
  labs(title = "Annual Mean Precipitation") + 
  facet_wrap(~NAME)
```