---
title: "Pretty Map"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("../setup.R")
```

```{r, warning=FALSE}
base_map()
```

Figure 2. A map view showing the extent of marine regions shown in blue: GBK = Georges Bank, WBN = Wilkinson Basin, WCS = Western Coastal Shelf, EMCS = Eastern Maine Coastal Shelf, JBN = Jordan Basin, GBN = Georges Basin. Global Historic Climate Network weather stations are shown with squares: A = Corinna, ME, B = Durham, NH, and C = Blue Hill Observatory, Milton MA. US Geological Survey river monitoring station at Androscoggin River near Auburn, Maine is shown with a triangle.  Current representations are EMCC = Eastern Maine Coastal Current, GBK = Georges Bank [?],
GS = Gulf Stream, NEC = Northeast Current, SCOT = Scotian Current, SLOPE = [?] and WMCC = Western Maine Coastal Current.


Base bathymetry from GEBCO Compilation Group (2023) GEBCO 2023 Grid (doi:10.5285/f98b053b-0cbc-6c23-e053-6c86abc0af7b). Marine regional boundaries adapted from https://www.marineregions.org/.  Androscoggin River watershed boundary adapted from U.S. Geological Survey, 2019, National Hydrography Dataset (ver. USGS National Hydrography Dataset Best Resolution (NHD) for Hydrologic Unit (HU) 4 - 2001 (published 20191002)), accessed August 23, 2023 at URL https://www.usgs.gov/national-hydrography/access-national-hydrography-products.
