---
title: "GOM-Series Buoys"
output: github_document
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
suppressPackageStartupMessages({
  library(cofbb)
})
source("../setup.R")

bb <- cofbb::get_bb("gom", form = 'sf')
regions = read_regions()
coast = rnaturalearth::ne_coastline(scale = "large", returnclass = 'sf') |>
  sf::st_crop(sf::st_bbox(bb))
```

## Buoy data aggregated to monthly means

5 buoys from across the Gulf of Maine were selected for inclusion.  A listing of these is saved in `data/buoy-listings.csv` and can be read as simple look-up table (lut).

```{r}
buoys = buoy_lut()
buoys
```

```{r, warning = FALSE}
ggplot(data = regions) +
  geom_sf(fill = NA) +
  geom_sf(data = coast, color = "blue") + 
  geom_point(data = buoys, aes(x = lon, y = lat), color = "orange") +
  geom_text(data = buoys, aes(x = lon, y = lat, 
                              label = id,
                              hjust = 0.5, 
                              vjust = 1.2))
```


## MET

### Fetch meteorological (met) data
Meteorological data for these buoys can be fetched using `fetch_buoy_met()`. This function downloads high temporal resolution data per buoy, and aggregates into monthly means, and saved to disk a simple table. Run this as needed to update data.

```
met <- lapply(buoys$id, fetch_buoy_met)
```

### Read and display met data

Read one or more buoy met data files using `read_buoy_met()`.  By default all buoys are read and boundinto one table.

```{r}
x <- read_buoy_met() |>
  dplyr::mutate(month = format(date, "%b"), .after = date) |>
  dplyr::group_by(station)

ggplot(data = x, aes(x = date, y = wind_speed.mean, color = station, shape = station)) +
  geom_line()
```

```{r}
ggplot(data = filter(x, month == 'Aug'), 
       aes(x = date, y = air_temperature.mean, color = station)) +
  geom_point() + 
  geom_smooth(method = "lm", se=FALSE) +
  labs(x = "year", y = 'Air Temp (C)', title = "August - mean monthly air temperature")
```


## CTD at depth (temp, salinity, sigma_t)

### Fetch CTD  data
CTD data for these buoys can be fetched using `fetch_buoy_ctd()`. This function downloads high temporal resolution data per buoy at various depths, and aggregates into monthly means, and saved to disk a simple table. Run this as needed to update data.

```
ctd <- lapply(buoys$id, fetch_buoy_ctd)
```

### Read and display CTD data

Read one or more buoy CTD data files using `read_buoy_ctd()`.  By default all buoys are read and bound into one table.

```{r}
x <- read_buoy_ctd() |>
  dplyr::mutate(depth = factor(as.integer(depth))) |>
  dplyr::mutate(month = format(date, "%b"), .after = date) |>
  dplyr::group_by(station)

ggplot(data = x, aes(x = date, y = temperature.mean, color = depth)) +
  scale_y_reverse()  + 
  geom_line() + 
  facet_wrap(~station)
```

## OPTICS at depth (solar_zenith_angle, Ed_PAR, chlorophyll)

### Fetch OPTICS data
OPTICS data for these buoys can be fetched using `fetch_buoy_optics()`. This function downloads high temporal resolution data per buoy at various depths, and aggregates into monthly means, and saved to disk a simple table. Run this as needed to update data. Not every buoy has optics data.

```
ctd <- lapply(buoys$id, fetch_buoy_optics)
```

### Read and display OPTICS data

Read one or more buoy CTD data files using `read_buoy_ctd()`.  By default all buoys are read and bound into one table.  Note it seems these measures are not at the surface but instead are at varying depths.

```{r}
x <- read_buoy_optics() |>
  dplyr::mutate(month = format(date, "%b"), .after = date) |>
  dplyr::mutate(water_depth = factor(water_depth)) |>
  dplyr::group_by(station)

count(x, station, water_depth)
```

```{r}
ggplot(data = x, aes(x = date, y = chlorophyll.mean, color = water_depth)) +
  scale_y_log10() + 
  geom_line() + 
  facet_wrap(~station)
```




## ACDP at depth (water_depth, depth, current_u, current_v)


### Fetch ACDP data

ACDP data for these buoys can be fetched using `fetch_buoy_acdp()`. This function downloads high temporal resolution data per buoy at various depths, aggregates into monthly means and yearly, and saved to disk simple tables. Run this as needed to update data. Not every buoy has ADCP data.

```
adcp <- lapply(buoys$id, fetch_buoy_adcp)
```

### Read and display ADCP data

Read one or more buoy ADCP data files using `read_buoy_adcp()`.  By default all buoys are read and bound into one table.  

```{r}
x <- read_buoy_adcp() |>
  dplyr::mutate(month = format(date, "%b"), .after = date) |>
  dplyr::mutate(water_depth = factor(water_depth)) |>
  dplyr::group_by(station)

count(x, station, water_depth)
```

```{r}
ggplot(data = x, aes(x = date, y = current_u.mean, color = water_depth)) +
  geom_line() + 
  facet_wrap(~station)
```

## RTSC at surface (current_speed, current_direction, temperature)

### Read and display RTSC data

Read one or more buoy RTSC data files using `read_buoy_rtsc()`.  By default all buoys are read and bound into one table.  

```{r}
x <- read_buoy_rtsc() |>
  dplyr::mutate(month = format(date, "%b"), .after = date) |>
  dplyr::mutate(water_depth = factor(depth)) |>
  dplyr::group_by(station)

count(x, station, depth)
```

```{r}
ggplot(data = x, aes(x = date, y = current_direction_u.mean)) +
  geom_line() + 
  geom_smooth(method = "lm", se=FALSE) +
  facet_wrap(~station)
```
