---
title: "Export"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning = FALSE, message = FALSE}
source("../setup.R")
```

Each data set function suite has it's own export functionality (by year, month and possibly both) which yields a wide-format table with a leading date column.  A single function, `export()` will export all of the datasets into one table (by year or by month). These may be subsequently read into R using the `read_export()` function.  Below we select columns that include "q75" (75th percentile) and filter for dates from 1950 until recent.  We standardize using the `scale` function which transforms each variable to have a mean of 0 and a standard deviation of 1.  

```{r}
x = read_export(by = 'year') |>
                dplyr::select(date, dplyr::contains("q25")) |>
                dplyr::filter(date >= as.Date("1950-01-01")) |>
                standardize_export() |>
  dplyr::glimpse()
```

We can plot a heat map for the above.
```{r, out.width = "100%", fig.height = 12}
plot_export(x, title = "Standardized Q25 by year", purge_empty = FALSE)
```


Other descriptive metrics can be shown.  This time we drop the buoy data and purge empty rows to improve readability.

```{r, out.width = "100%", fig.height = 12}
read_export(by = 'year') |>
                dplyr::select(date, dplyr::contains("q75") & !dplyr::starts_with("BUOY")) |>
                dplyr::filter(date >= as.Date("1950-01-01")) |>
                standardize_export() |>
  plot_export(title = "Standardized Q75 by year", purge = TRUE)
```


And now with buoys only, note that we filter for dates after 2000.

```{r, out.width = "100%", fig.height = 12}
read_export(by = 'year') |>
                dplyr::select(date, dplyr::starts_with("BUOY") & dplyr::contains("median") ) |>
                dplyr::filter(date >= as.Date("2000-01-01")) |>
                standardize_export() |>
  plot_export(title = "Standardized median by year", purge = FALSE)
```


