---
title: "GOM-series Input Correlation Test"
output: github_document
---


```{r message=FALSE, warning=FALSE}
source("../setup.R")
```

```{r}
x <- read_export(by = "year")
```

Use the function `network(x, include = c(vars))` where x is the table from `read_export()` and include is a character vector of what variables from x to include in the plot (these can be individual variable names or keywords). 

## Climate Indices

Each index is grouped with itself;
NAO seems to be the most spread out;
NAO and AMO are negatively correlated

```{r climate_network}
network(x, include=c("nao", "gsi", "amo"))
```

## PCI and the HAB Index



```{r pci_hab_network, message=FALSE, warning=FALSE}
network(x, include=c("pci", "hab"))
```

## USGS

```{r usgs_network, message=FALSE, warning=FALSE}
network(x, include=c("usgs"))
```

## SST

The coastal shelf minimums seperate from the pack;
Some Georges Bank vars are on their own too

```{r sst_network, message=FALSE, warning=FALSE}
network(x, include=c("sst"))
```

## SST + HAB

All negative relationships between both HAB indices and the SST vars

```{r sst_hab_network, message=FALSE, warning=FALSE}
network(x, include=c("sst", "hab"))
```

## Chlorophyll

Notice I had to use "chlor." to exclude the buoy vars with "chlorophyll"

```{r chlor_network, message=FALSE, warning=FALSE}
network(x, include=c("chlor."))
```

## Try them all?

Things get busy...

```{r all_network, message=FALSE, warning=FALSE}
network(x, include=c("nao", "gsi", "amo", "sst", "hab", "pci", "usgs", "chlor."))
```

