---
title: "GOM-series Input Correlation Test"
output: github_document
---


```{r message=FALSE, warning=FALSE}
source("../setup.R")
```

```{r}
x <- read_export(by = "year")
```

## Test the Selected Analysis Variables 

Use the function `analysis_vars()` to read in the subset of variables and supply them ti the include argument of `network()`

### Median

```{r}
vars <- analysis_vars(treatment = "median")

vars
```

```{r median_network, warning = FALSE, message = FALSE}
x |>
  network(include = vars)
```

### q25

```{r q25_network, warning = FALSE, message = FALSE}
network(x, include = analysis_vars(treatment = "q25"))
```



### q75

```{r q75_network, warning = FALSE, message = FALSE}
network(x, include = analysis_vars(treatment = "q75"))
```


Initial correlation plots:

## Climate Indices

Each index is grouped with itself;
NAO seems to be the most spread out;
NAO and AMO are negatively correlated

```{r climate_network}
x |>
    dplyr::select(dplyr::contains(c("nao", "gsi", "amo"))) |>
    corrr::correlate() |>
    corrr::network_plot(colours = c("skyblue1", "white", "indianred2"))
```

## PCI and the HAB Index



```{r pci_hab_network, echo=FALSE, message=FALSE, warning=FALSE}
x |>
    dplyr::select(dplyr::contains(c("pci", "hab"))) |>
    corrr::correlate() |>
    corrr::network_plot(colours = c("skyblue1", "white", "indianred2"))

```

## USGS

```{r usgs_network, echo=FALSE, message=FALSE, warning=FALSE}
x |>
    dplyr::select(dplyr::contains(c("usgs"))) |>
    corrr::correlate() |>
    corrr::network_plot(colours = c("skyblue1", "white", "indianred2"))

```

## SST

The coastal shelf minimums seperate from the pack;
Some Georges Bank vars are on their own too

```{r sst_network, echo=FALSE, message=FALSE, warning=FALSE}
x |>
    dplyr::select(dplyr::contains(c("sst"))) |>
    corrr::correlate() |>
    corrr::network_plot(colours = c("skyblue1", "white", "indianred2"))

```

## SST + HAB

All negative relationships between both HAB indices and the SST vars

```{r sst_hab_network, echo=FALSE, message=FALSE, warning=FALSE}
x |>
    dplyr::select(dplyr::contains(c("sst", "hab"))) |>
    corrr::correlate() |>
    corrr::network_plot(colours = c("skyblue1", "white", "indianred2"))

```

## Chlorophyll

Notice I had to use "chlor." to exclude the buoy vars with "chlorophyll"

```{r chlor_network, echo=FALSE, message=FALSE, warning=FALSE}
x |>
    dplyr::select(dplyr::contains(c("chlor."))) |>
    corrr::correlate() |>
    corrr::network_plot(colours = c("skyblue1", "white", "indianred2"))

```

## Try them all?

Things get busy...

```{r all_network, echo=FALSE, message=FALSE, warning=FALSE}
x |>
    dplyr::select(dplyr::contains(c("nao", "gsi", "amo", "sst", "hab", "pci", "usgs", "chlor."))) |>
    corrr::correlate() |>
    corrr::network_plot(colours = c("skyblue1", "white", "indianred2"))

```

